---
title: "20-explore"
output: html_notebook
---

## Overview

For the operation in the future, having a overview can help us understand the dataset easily.

```{r overview}
dim(USvideos)
str(USvideos)
```


## Assert outlier

Now we need to make sure is there any outlier or mistake in the dataset.


### Assert category_id

First, test the column called "category_id". There are 43 categories, therefore the values in the column should not be bigger than 43 or smaller than 1.

```{r assert category}
assert(data = USvideos, in_set(1, 43, allow.na = FALSE), category_id) 
```

There are 5 rows have *NA* in this column, we can just remove them later.

### Assert numerical columns

For the numerical columns in the dataset, based on the reality, all of them should be positive.

```{r assert positive number}
assert(data = USvideos, within_bounds(lower.bound = 0, upper.bound = Inf, allow.na = FALSE), views)
assert(data = USvideos, within_bounds(lower.bound = 0,upper.bound = Inf, allow.na = FALSE), likes)
assert(data = USvideos, within_bounds(lower.bound = 0, upper.bound = Inf, allow.na = FALSE), dislikes)
assert(data = USvideos, within_bounds(lower.bound = 0, upper.bound = Inf, allow.na = FALSE), comment_count)
```

Fortunately, all of the numbers are positive. There is no mistake.

### Aassert logical columns

And for the logical columns, all of the values should be TRUE or FALSE.

```{r assert logical}
assert(data = USvideos, in_set(TRUE, FALSE, allow.na = FALSE), comments_disabled)
assert(data = USvideos, in_set(TRUE, FALSE, allow.na = FALSE), ratings_disabled)
assert(data = USvideos, in_set(TRUE, FALSE, allow.na = FALSE), video_error_or_removed) 
```

And there is no error too.


## Clean the data. 

### Remove *NA*

Because there are only several observations with NA values, we can just remove all of the rows which have NA value.
```{r remove NA}
USvideos_NNA <- as.data.frame(na.omit(USvideos))
USvideos_NNA
```

### Convert date column

Then we need to convert the column called "trending_date" with character type to normal date format in "lubridate" package.

```{r comvert to lubridate}
USvideos_NNA <- USvideos_NNA %>%
  mutate(trending_date = ydm(trending_date))
```


## Result overview

Now let's look through the structure of dataset again.

```{r result overview}
str(USvideos_NNA)
```

